### 2014-10-31 contract-language.tex
edits

Our goal is to map application-level consistency constraints on
operations to appropriate store-level consistency guarantees capable
of satisfying these constraints.  The ability to express both of them as
contracts in our contract language lets us compare and determine if
contract ($\cv_{op}$) of an operation ($op$) is weak enough to be
satisfied under a store consistency level identified by the contract
$\cv_{st}$. Towards this end, we define a binary \emph{weaker than}
relation for our contract language as following:

@@

If the contract ($\cv_{op}$) of an operation ($op$) is \emph{weaker
than} a store contract ($\cv_{st}$), then constraints expressed by the
former are implied by guarantees provided by the later. Due to the
completeness of first-order logic\cite{completeness}, this means that
any well-formed execution ($\E$) that satisfies $\cv_{st}$ (i.e.,
$\E\models \cv_{st}$) also satisfies $\cv_{op}$ (i.e., $\E \models
\cv_{op}$). Consequently, it is safe to execute operation $op$ under
store consistency level captured by $\cv_{st}$.

@@

We model the system as a tuple $\E;\Sigma$, where the execution $\E$
captures the data store's current state of the execution, and
$\Sigma$, the session soup, is the set of concurrent client sessions
interacting with the store. A session is modeled simply as a sequence
of replicated data type operations ($op$), tagged with the consistency
class ($\tau$) of their contracts (as determined by the contract
classification scheme). We assume a rewrite relation of form:

@@

\noindent on the system state. The relation captures the progress of
the execution (from $\E$ to $\E'$)  due to the application of an
operation ($op$) from one of the sessions in $\Sigma$ to a replica,
generating a new effect ($\eff$). If the resultant execution ($\E'$)
satisfies the store contract ($\cv(\tau)$) of $\tau$ (i.e., $\E
\models \cv(\tau)$), then we say that the store has \emph{enforced} the
contract $\cv(\tau)$ in the execution $\E'$.

@@

\begin{theorem}[Soundness of Contract Enforcement]
\label{lem:core-preservation}
Let $\cv$ be a well-formed contract of a replicated data type operation $op$,
and let $\tau$ denote the consistency class of $\cv$ as determined by
the contract classification scheme. Forall well-formed execution
states $\E$, $\E'$ such that
$\auxred{} {\E;\langle (op,\tau);\,\sigma \rangle \pll \Sigma} {\eff}
 {\E';\langle \sigma \rangle \pll \Sigma}$
if $\E' \models [\eff/\cureff]\,\cv(\tau)$, then $\E' \models [\eff/\cureff]\,\cv$
\end{theorem}

>>>>

Our goal is to map application-level consistency constraints on
operations to appropriate store-level consistency guarantees capable
of satisfying these constraints.  The ability to express both of them as
contracts in our contract language lets us compare and determine if
contract ($\cv_{op}$) of an operation ($\mathit{op}$) is weak enough to be
satisfied under a store consistency level identified by the contract
$\cv_{st}$. Towards this end, we define a binary \emph{weaker than}
relation for our contract language as following:

@@

If the contract ($\cv_{op}$) of an operation ($\mathit{op}$) is \emph{weaker
than} a store contract ($\cv_{st}$), then constraints expressed by the
former are implied by guarantees provided by the later. Due to the
completeness of first-order logic\cite{completeness}, this means that
any well-formed execution ($\E$) that satisfies $\cv_{st}$ (i.e.,
$\E\models \cv_{st}$) also satisfies $\cv_{op}$ (i.e., $\E \models
\cv_{op}$). Consequently, it is safe to execute operation $\mathit{op}$ under
store consistency level captured by $\cv_{st}$.

@@

We model the system as a tuple $\E;\Sigma$, where the execution $\E$
captures the data store's current state of the execution, and
$\Sigma$, the session soup, is the set of concurrent client sessions
interacting with the store. A session is modeled simply as a sequence
of replicated data type operations ($\mathit{op}$), tagged with the consistency
class ($\tau$) of their contracts (as determined by the contract
classification scheme). We assume a rewrite relation of form:

@@

\noindent on the system state. The relation captures the progress of
the execution (from $\E$ to $\E'$)  due to the successful completion
of a client operation ($\mathit{op}$) from one of the sessions in $\Sigma$,
generating a new effect ($\eff$). If the resultant execution ($\E'$)
satisfies the store contract ($\cv(\tau)$) of $\tau$ (i.e., $\E
\models \cv(\tau)$), then we say that the store has \emph{enforced} the
contract $\cv(\tau)$ in the execution $\E'$.

@@

\begin{theorem}[Soundness of Contract Enforcement]
\label{lem:core-preservation}
Let $\cv$ be a well-formed contract of a replicated data type operation $\mathit{op}$,
and let $\tau$ denote the consistency class of $\cv$ as determined by
the contract classification scheme. Forall well-formed execution
states $\E$, $\E'$ such that
$\auxred{} {\E;\langle (op,\tau);\,\sigma \rangle \pll \Sigma} {\eff}
 {\E';\langle \sigma \rangle \pll \Sigma}$
if $\E' \models [\eff/\cureff]\,\cv(\tau)$, then $\E' \models [\eff/\cureff]\,\cv$
\end{theorem}

### 2014-10-31 transactions.tex
edits

We assume that operations not part of any transaction belong to their own
unique transaction. While transactions may have varying isolation
guarantees, we make the standard assumption that all transactions provide
atomicity. Hence, we include the following axiom in $\Delta$: $\small \forall
a,b,c.~\txn{a}{b,c} ~\wedge~ \sameobj{b}{c} ~\wedge~ \vis{b}{a} \Rightarrow
\vis{c}{a}$. The semantics of this contract is illustrated represented in
Figure~\ref{fig:txn_atomicity}.

@@

In order to illustrate the utility of declarative reasoning for transactions,
let us extend our running bank account example to use two accounts (objects) --
current ($c$) and savings ($s$). Each account provides operations
\cf{withdraw}, \cf{deposit} and \cf{getBalance}, with the same contracts as
defined previously. We consider two transactions -- \cf{save(amt)}, which
transfers \cf{amt} from current to savings, and \cf{totalBalance}, which
returns the sum of the balances of individual accounts. The \name code for the
transactions is given below:
\vspace{-1em}

>>>>

We assume that operations not part of any transaction belong to their own
unique transaction. While transactions may have varying isolation
guarantees, we make the standard assumption that all transactions provide
atomicity. Hence, we include the following axiom in $\Delta$: \[\small \forall
a,b,c.~\txn{a}{b,c} ~\wedge~ \sameobj{b}{c} ~\wedge~ \vis{b}{a} \Rightarrow
\vis{c}{a}\] \noindent The semantics of this contract is illustrated in
Figure~\ref{fig:txn_atomicity}.

@@

In order to illustrate the utility of declarative reasoning for transactions,
consider an extension of our running example to use two accounts (objects) --
current ($c$) and savings ($s$). Each account provides operations
\cf{withdraw}, \cf{deposit} and \cf{getBalance}, with the same contracts as
defined previously. We consider two transactions -- \cf{save(amt)}, which
transfers \cf{amt} from current to savings, and \cf{totalBalance}, which
returns the sum of the balances of individual accounts. The \name code for the
transactions is given below:
\vspace{-1em}
